### 2024-12-03 paper/paper.qmd
updated paper

---
title: "Family Income and Educational Returns in China, 2010-2020: Higher Returns Favor the Wealthy, but Gaps Hold Steady"
author: 
  - Yisu Hou
thanks: "Code and data are available at: [https://github.com/YisuHou1/Educational_Returns_Analysis](https://github.com/YisuHou1/Educational_Returns_Analysis)."
date: today
date-format: long
abstract: "Economic returns to education are a key driver of socioeconomic mobility, yet they are not distributed equally across all income groups. Using data from the China Family Panel Studies (CFPS) between 2010 and 2020, this paper analyzes how family income influences the financial benefits individuals gain from education. The analysis reveals that individuals from wealthier families consistently enjoy significantly higher income returns from education compared to their lower-income counterparts, though this gap did not widen over the decade studied. These findings highlight persistent inequality in education's economic impact, emphasizing the need for policies that ensure education serves as a more effective equalizer in society."
format: pdf
header-includes:
  - \usepackage{float}
  - \usepackage{placeins}
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(caret)
library(glmnet)
library(nnet)
library(stringr)
library(gt)
library(broom)
library(knitr)
library(MatchIt)
library(arrow)
```

# Introduction

Since the economic reform in China in 1978, China’s GDP per capita has increased more than 29 times [@NBSC2020]. However, this period of rapid economic growth has been accompanied by greater socioeconomic inequality. The increased stratification, along with a number of China specific policies, has pushed the GINI coefficient close to 0.5, and various studies have demonstrated that the majority of China’s wealth is concentrated in the hands of a few [@XieJin2015].

In examining the factors contributing to China’s increasing inequality, a multitude of social scientists highlight China’s education system. The correlation between socioeconomic status and education has been widely scrutinized. According to Esping-Andersen, "in modern capitalist society, the two main exclusionary devices through which the bourgeoisie structures and perpetuates itself as a class, are first, those surrounding the institutions of property, and second, academic or professional qualifications and credentials." [@EspingAndersen1982]

In China, educational attainment directly corresponds to future earnings. Research by Byron and Manaloto showed an approximate 4 percent increase in earnings for each year of schooling [@Byron1990], and Wu and Xie identified a rate of 4.7 percent [@Wu2003]. Jansen and Wu documented a significant rising trend of returns on education during China’s economic reform period from 1978 to 2005 [@Jansen2012]. Concurrently, China significantly expanded its higher education sector, raising the gross enrollment rate from 9.8 percent in 1998 to 24.2 percent in 2009 [@Wang2011]. As of 2022, the enrollment rate stands at 59.7 percent, indicating that China has fully transitioned from elite to mass higher education. Although increased access to higher education and higher returns on education are generally believed to foster social mobility, China’s inequality has worsened. To elucidate this paradoxical outcome, researchers found that the new educational opportunities are predominantly accessed by privileged groups, such as individuals with higher social status, males, and urban populations [@Ou2019].

Previous research has also investigated the reasons behind the educational gap between high and low-income individuals. In the 20th century, many theorists proposed theories examining the link between socioeconomic status and educational achievement. Some suggest that income-based segregation places low-income children in overcrowded classrooms with less adequately trained educators. Others argue that teachers may harbor negative attitudes towards low-income children, anticipating they will perform poorly academically, thereby creating a self-fulfilling prophecy [@Bond1981]. More recently, Duncan and Murnane suggested that the disparity in child enrichment expenditure influences academic performance, giving children from high-income families an advantage [@Duncan2014].

However, while the relationship between financial status and educational attainment is relatively well established, there is a dearth of research, especially in China, to explore the link between family financial status and the rate of return from education. This is an essential connection, as it signifies how effectively education translates into its ultimate outcome: income. The aim of this research is to fill this knowledge gap by examining whether individuals from different socioeconomic backgrounds in the Chinese population receive the same level of returns from education.

The estimand of this investigation is the disparities in educational returns among individuals from different family wealth quartiles during the decade 2020 to 2020, measured in the percent income increase per year of education received.

To arrive at a conclusion, national family and individual data from China Family Panel Studies [@cfps] from the year 2010 to 2020 were extracted. For the data in each CFPS iteration, Propensity Score Matching (PSM) was utilized to balance the control variables across different family wealth quartiles, and separate polynomial regressions were fit to identify the disparities of educational returns across quartiles. It was found that throughout the decade, individuals from wealthier families possessed a significant advantage over individuals from less wealthy families, with disparities reaching more than 3% income increase per year of education between some quartiles.

The body of the paper contains the data, model, results, and discussion sections. In order, @sec-data contains an overview of the data, the variables used, and the programming tools used in my analysis and interpretation of the data. @sec-model describes the model in detail and provides justification for it. @sec-results contains a run-down and explanation of the results and findings from the model, and @sec-discussion contains a discussion of the implications and limitations of the result, among others. The appendices contain a detailed examination of the methodologies of CFPS, as well as aspects of the data and analysis excluded from the body.

# Data {#sec-data}

## Overview

I used the statistical programming language R [@R] for the graphing, analysis and presentation of the project as a whole. Caret [@caret], glmnet [@glmnet; @glmnet2; @glmnet3], nnet [@nnet], tidyverse [@tidyverse], car [@car] MatchIt [@matchit] and arrow [@arrow] were used directly in the analysis of the data. Stringr [@stringr], dplyr [@dplyr], gt [@gt], knitr [@knitr], and styler [@styler] were used in the presentation and/or styling of the data, graphs, and paper as a whole.

All data utilized in this study were sourced from various iterations of the China Family Community Survey [@cfps]. The CFPS is a nationwide community survey funded by 985 Program of Peking University and carried out by the Institute of Social Science Survey of Peking University. CFPS encompasses approximately 15,000 families across 25 provinces and aims to tracks them permanently since 2010. Its main objective is to gather data across three dimensions: individuals, households, and communities. To achieve this, CFPS comprises four sub-surveys that thoroughly investigates families, adults, and children, gathering data for thousands of variables. Given its expansive and representative sample and a wide range of variables, CFPS has become a staple dataset for research in Chinese social sciences. Alternative data sources for this study, such as China General Social Survey (CGSS) and China Social Survey (CSS) are also reliable but do not provide the same width of variables, especially concerning family units. For the purposes of this research, data was extracted from CFPS for the years 2010, 2012, 2014, 2016, 2018, and 2020.

After cleaning the data to split results by candidate, removing polls with N/A values and ratings lower than 2.5, and further selecting polls polling likely voters as opposed to the whole population, we are left with 746 data points with the variables pollster, has_sponsor, numeric_grade, pollscore, transparency_score, sample_size, end_date, state, candidate_name, pct (support rate in percentage), and cycle (election cycle, i.e. 2020/2024).

## Measurement

## Outcome variable

The variables presented in this section are variables in the polynomial regression model to identify returns to education, which is a coefficient of the model. This will be specified in @sec-model.

The outcome variable of the model is the natural logarithm of individuals' hourly wages. This variable is not directly available in CFPS. It was calculated by dividing individuals' yearly work hours from their yearly income. @fig-outcome shows its distribution in the aggregated survey data from all CFPS iterations between 2010 and 2020.

```{r}
#| echo: false
#| message: false
#| warning: false

data_2020 <- read_parquet(here::here("data/02-analysis_data/2020_analysis_data.parquet"))
data_2018 <- read_parquet(here::here("data/02-analysis_data/2018_analysis_data.parquet"))
data_2016 <- read_parquet(here::here("data/02-analysis_data/2016_analysis_data.parquet"))
data_2014 <- read_parquet(here::here("data/02-analysis_data/2014_analysis_data.parquet"))
data_2012 <- read_parquet(here::here("data/02-analysis_data/2012_analysis_data.parquet"))
data_2010 <- read_parquet(here::here("data/02-analysis_data/2010_analysis_data.parquet"))

# Make a combined dataset that includes all survey iterations
all_combined <- bind_rows(
  data_2020, data_2018, data_2016, data_2014, data_2012, data_2010
)
```

```{r}
#| label: fig-outcome
#| fig-cap: Natural Log of Individual Hourly Wage, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

# Outcome variable: ln hourly wage
ggplot(all_combined, aes(x = ln_hourly_wage)) +
  geom_histogram(
    fill = "#2c7bb6", # Aesthetic fill color (steel blue)
    color = "white", # Border color for bins
    alpha = 1 # Transparency of the fill
  ) +
  labs(
    x = "ln Hourly Wage",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  ) +
  # Add a vertical line for the mean percentage support
  geom_vline(
    aes(xintercept = mean(ln_hourly_wage, na.rm = TRUE)),
    color = "#d73027", # Aesthetic color (red)
    linetype = "dashed",
    size = 1
  ) +
  # Annotate the mean value on the plot
  annotate(
    "text",
    x = mean(all_combined$ln_hourly_wage, na.rm = TRUE) - 3, # Position text slightly to the right of the mean line
    y = Inf, # Position text at the top of the plot
    label = paste("Mean:", round(mean(all_combined$ln_hourly_wage, na.rm = TRUE), 1)),
    vjust = 2, # Vertical adjustment
    color = "#d73027",
    size = 5,
    fontface = "bold"
  )
```

\FloatBarrier

As displayed by @fig-outcome, the distribution of the natural log of hourly wage is close to a normal distribution, with a mean of 2.2 and most data points between 0 and 5. However, the range of values extends from around -9 to 12, representing individuals with extremely low and high income.

## Predictor variables

### Years of Education Received

Years of education is selected as the primary predictor in the model to identify economic returns to education. The theoretical framework for the model will be explained in @sec-model.

```{r}
#| label: fig-edu
#| fig-cap: Years of Education Received, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = years_of_education)) +
  geom_histogram(
    binwidth = 1,
    fill = "#2c7bb6", # Aesthetic fill color (steel blue)
    color = "white", # Border color for bins
    alpha = 1 # Transparency of the fill
  ) +
  labs(
    x = "Years of Education",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  ) +
  # Add a vertical line for the mean percentage support
  geom_vline(
    aes(xintercept = mean(years_of_education, na.rm = TRUE)),
    color = "#d73027", # Aesthetic color (red)
    linetype = "dashed",
    size = 1
  ) +
  # Annotate the mean value on the plot
  annotate(
    "text",
    x = mean(all_combined$years_of_education, na.rm = TRUE) + 2.5, # Position text slightly to the right of the mean line
    y = Inf, # Position text at the top of the plot
    label = paste("Mean:", round(mean(all_combined$years_of_education, na.rm = TRUE), 1)),
    vjust = 2, # Vertical adjustment
    color = "#d73027",
    size = 5,
    fontface = "bold"
  )
```

\FloatBarrier

As shown in @fig-edu, most individuals in the aggregate dataset take very specific values, such as 9, 12, and 15. This is because most iterations of the CFPS record education received as the highest level of education completed. During the data cleaning process, they were translated to their corresponding number of years. For example, an individual who completed high school receives a 12 for years of education. I am aware that this could be highly inaccurate, as individuals may skip or repeat a grade. This is a limitation of the study that will be fully discussed in @sec-discussion.

### Gender

In light of the burgeoning prominence of the gender wage gap as a global sociological concern in recent decades, extensive research has been undertaken to discern the connection between gender and income. Scholars discovered that Chinese females earned 84.4 percent of male wages in 1988 and 82.5 percent in 1995 [@GustafssonLi2000]. Zhang identified a 38.23 percent wage gap using 2004 data, demonstrating an exacerbation of the issue over time [@Zhang2006].

Beyond merely influencing wage, studies have found a direct bearing of gender on educational returns, particularly in China. Gustafsson and Li found that in 1988, a college diploma raised a woman’s income by 10.2 percent and a man’s income by 8.9 percent compared to high school graduates [@GustafssonLi2000]. In 1995, the values changed to 20.8 percent for women 15.5 percent for men. A more recent study questioned the real magnitude of the gender-based difference in educational returns, as Chinese couples often pool income together, and many females have access to their spouse’s income [@Hannum2013]. Nevertheless, the role of gender as a determinant influencing wage and educational returns is well-established. Thus, it was included as a predictor.

```{r}
#| label: fig-gender
#| fig-cap: Gender of Respondent, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = factor(gender, levels = c(0, 1), labels = c("Female", "Male")))) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Gender",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

@fig-gender shows that there were significantly more male CFPS respondents than female respondents throughout 2010-2020.

### Work Experience

Work experience is another key variable in the model for educational returns. CFPS does not directly provide data for work experience, so potential work experience was created as a substitute.

Potential work experience was deduced with an individual’s age and their years of schooling. Established norms in labor economics prescribe the following age thresholds for labor market entry based on educational attainment: 17 for high-school dropouts, 19 for high-school graduates, 21 for high-school graduates with some college, and 23 for college graduates [@Borjas2003; @OttavianoPeri2012; @EdoRapoport2019]. Thus, potential work experience was calculated by subtracting the corresponding value from an individual's age. This assumes that individuals directly enter the labor market after completing their education, which is a strong assumption.

```{r}
#| label: fig-exp
#| fig-cap: Potential Years of Work Experience, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

# Potential work experience
ggplot(all_combined, aes(x = pot_work_years)) +
  geom_histogram(
    fill = "#2c7bb6", # Aesthetic fill color (steel blue)
    color = "white", # Border color for bins
    alpha = 1 # Transparency of the fill
  ) +
  labs(
    x = "Potential Years of Work Experience",
    y = "Frequency"
  ) +
  coord_cartesian(xlim = c(0, 60)) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  ) +
  # Add a vertical line for the mean percentage support
  geom_vline(
    aes(xintercept = mean(pot_work_years, na.rm = TRUE)),
    color = "#d73027", # Aesthetic color (red)
    linetype = "dashed",
    size = 1
  ) +
  # Annotate the mean value on the plot
  annotate(
    "text",
    x = mean(all_combined$pot_work_years, na.rm = TRUE) + 6, # Position text slightly to the right of the mean line
    y = Inf, # Position text at the top of the plot
    label = paste("Mean:", round(mean(all_combined$pot_work_years, na.rm = TRUE), 1)),
    vjust = 2, # Vertical adjustment
    color = "#d73027",
    size = 5,
    fontface = "bold"
  )
```

\FloatBarrier

@fig-exp shows that potential work experience is a right-skewed distribution with a mean of 17.8. This is expected, as it is unlikely for individuals to work for more than 40 years.

### City Hukou Status

The Hukou system in China, a household registration mechanism, classifies individuals into either rural or urban demographics based on their birth registration locale. Individuals with urban Hukou and live in major cities enjoy the lion share of quality education and opportunities to pursue higher education. According to Tam and Jiang, during the recent period of higher education expansion, urban students’ access to higher education rose sharply while rural students’ access stagnated, showing the growing advantage of urban students due to unequal distribution of educational resources [@Tam2015]. The scholars explained the widening gap by pointing to the expansion of vocational education, which were accessible to low-performing urban students who would not have had the opportunity to pursue higher education prior to the expansion. However, rural students were not provided with equal opportunities in vocational education.

How the Hukou system affects educational attainment is a complex topic that must be examined together with other phenomena in China. Local governments are responsible for educational institutions within its region, but rural governments only have a minimal proportion of the capital available to large cities; top universities are disproportionally located in major cities and are more available to individuals possessing a Hukou of those cities; migrant workers, or individuals who live in cities but do not hold a Hukou of the city they live in, do not have access to public schooling in the city and may only attend migrant schools with worse conditions and educational quality [@Wang2007].

Therefore, it is no surprise that urban Hukou holders have higher wages. Using data from 2005, scholars found a rural Hukou worker’s wage per hour is only 64 percent of an urban Hukou worker’s [@Xing2013].

```{r}
#| label: fig-hukou
#| fig-cap: Hukou Status of Respondent, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = factor(city_hukou, levels = c(0, 1), labels = c("Rural", "Urban")))) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Hukou Categorization",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

@fig-hukou shows that there are a greater number of rural Hukou holders than urban Hukou holders among CFPS respondents. This accurately reflects the overall demographics of China, where the rural population is larger than the urban population.

### Marital Status

Studies have shown the magnifying effect of marriage on gendered income gaps, stemming from gender-based roles and expectations within familial settings. Wang and Wang demonstrated that unmarried Chinese men and women in 2015 had an average monthly income gap of 458.612 RMB, and married couples had a gap of 1398.80 RMB [@WangWang2020]. Married couples with children under 18 had an income gap that is more than doubled compared to those without children. Mu and Xie explained the after-marriage changes by addressing that fathers acquire a confidence boost after the birth of their children and dedicate themselves in the pursuit for higher wages and a successful career [@Mu2016]. On the other hand, mothers take the child-rearing role, accumulate less experience in the labor force, and choose less rigorous jobs to reserve energy for household work.

```{r}
#| label: fig-marry
#| fig-cap: Marital Status of Respondent, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = factor(is_married, levels = c(0, 1), labels = c("Unmarried", "Married")))) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Marital Status",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

As displayed in @fig-marry, most respondents of CFPS are married.

### Communist Party Membership

Membership in the Communist Party of China (CCP) is frequently included in income related research within China [@GustafssonLi2000; @Hannum2013; @Mu2016]. Being a member of the CCP grants the opportunity for an individual to work in the government positions or state-owned enterprises that provide secure and well-paid jobs [@Dickson2014]. Moreover, party status is also related to the level of education, as the recruitment strategy of the CCP targets elite groups such as college and graduate students. This study followed established protocols and included CCP membership as a predictor, as its unique socio-political status in China necessitates its inclusion.

```{r}
#| label: fig-ccp
#| fig-cap: CCP Membership of Respondent, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = factor(is_party, levels = c(0, 1), labels = c("Not CCP Member", "CCP Member")))) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Party Membership",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

@fig-ccp reflects that CCP members are only make up a small proportion of the Chinese population.

### Eastern Provinces

China's geographical East, encompassing 12 economically advanced provinces, is characterized by notable income disparities when compared to Central and Western regions. Scholars found that in 2005, Eastern populations not only had significantly greater income but also enjoyed doubled returns to education compared to other regions [@Xing2013]. Though the values shifted towards the national mean in 2011, significant differences still existed. Recent research endeavors on income and educational returns have increasingly incorporated the geographical East as a control variable, acknowledging its substantial influence on the variables of interest [@Dong2021].

```{r}
#| label: fig-east
#| fig-cap: Location of Residence, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = factor(is_east, levels = c(0, 1), labels = c("Other Provinces", "Eastern Provinces")))) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Area of Residency",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

According to @fig-east, there exists a relatively balanced split between respondents residing in Eastern provinces and those living in other provinces. More individuals lived in non-Eastern provinces, counting to more than 20,000.

### Family Income Quartile

Family income quartile is not a variable directly incorporated in the regression model. Rather, data was split into subsets according to family income quartiles and regressed to find the educational returns of each quartile in comparison with one another. This is the focal variable of the study. Within the CFPS's family economics sub-survey, there exists a variable listing family income quartiles. For this study, family economics datasets were merged with the adult individuals dataset of the same CFPS iteration by matching personal IDs, thus deriving a corresponding family income quartile value for each adult participant.

```{r}
#| label: fig-quartiles
#| fig-cap: Family Income Quartiles, CFPS2010 to CFPS2020
#| echo: false
#| message: false
#| warning: false

ggplot(all_combined, aes(x = income_quartile)) +
  geom_bar(fill = "#1f78b4", color = "white", alpha = 1) +
  labs(
    x = "Family Income Quartile",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 14) + # Minimal theme with increased base font size
  theme(
    plot.title = element_text(
      size = 16,
      face = "bold",
      hjust = 0.5, # Center the title
      margin = margin(b = 10) # Add space below the title
    ),
    plot.subtitle = element_text(
      size = 12,
      hjust = 0.5,
      margin = margin(b = 20) # Add space below the subtitle
    ),
    axis.title = element_text(face = "bold"), # Bold axis titles
    axis.text = element_text(color = "gray30"), # Dark gray axis text
    panel.grid.major = element_line(color = "gray80"), # Light gray major grid lines
    panel.grid.minor = element_blank() # Remove minor grid lines
  )
```

\FloatBarrier

The distribution shown in @fig-quartiles is counter intuitive, as there are more individuals who belong in upper quartiles. This does not suggest that CFPS uses an unrepresentative sample. The data of individuals in upper family income quartiles are relatively more complete, thus more data points are kept after the data cleaning process. Despite the relative lack of data for the lower quartiles, the number of data points is still sufficient for this study.

# Model {#sec-model}

The goal of my modelling strategy is twofold. First, I must split the data from each CFPS iteration into family income quartiles and adjust them to be directly comparable with one another. Second, I wish to calculate each quartile's returns to education, in terms of percent income increase per year of education, to complete the comparison.

To accomplish the first task, Propensity Score Matching was utilized.

Historically, research on returns to education primarily relied on directly calculating the rate of return using the Mincer Earnings Function. They incorporated control variables into the regression to account for their influence on income [@Byron1990; @Zhang2006]. When desiring to make group comparisons, scholars generally subdivided data based on specific variables of interest such as gender, location, and industry. They then performed direct regressions to contrast the coefficients [@Gustafsson2000, @Mu2016].

While multivariate regression effectively captures the impact of control variables, its functionality hinges on the linear relationships between the outcome and covariates. Dehejia and Wahba revealed that regressing groups with significant variances in the mean values of covariates could yield misleading results for the variable of interest, potentially overstating the differences between groups [@Dehejia2002]. Propensity Score Matching (PSM) addresses this issue by isolating the treatment effect. Therefore, this study employed a combined method: first balancing covariates using PSM, then executing regressions to adjust for residual differences.

For this study, data from each CFPS iteration were divided into four subsets based on family income quartiles. PSM was then applied to match each subset with others, resulting in six pairs of subsets that exhibited different treatment variable values (family income quartiles) but had similar means for control variables. The matching was conducted using the nearest-neighbor approach with $caliper = 0.1sd$, which is more strict compared to the generally recommended 0.2 threshold for caliper width [@Austin2011]. This choice was intentional, as reducing the caliper width improved the covariate balance without significantly reducing the number of data points. Given that lower family income quartiles had fewer data points, match ratios were adjusted based on the counts within the quartiles being matched.

Table 2 provides a summary of balance for the match between the Quartile 1 and the Quartile 2 subset of CFPS 2020. The match was executed using caliper = 0.1 and a ratio of 2, matching 1041 Quartile 2 instances out of 1324 total instances with all 541 Quartile 1 instances. This matching procedure was repeated 6 times to compare each quartile with every other quartile for all 6 CFPS iterations from 2010 to 2020. The standard mean differences of matches using caliper = 0.2 and 0.5 were included as a comparison. The caliper of 0.1 returned the best standardized mean difference of 0.0139 between propensity scores of Q1 and Q2. Caliper widths of 0.2 and 0.5 returned 0.0338 and 0.0795, respectively.

TABLE HERE IF TIME

For the second task, a multivariate regression model was created using the theoretical framework of the Mincer Earnings Function [@Mincer1974]. The function depicts wage as a function of schooling and work experience. Its original form is:
$$ln(w) = f(s, x) = ln(w_0) +\rho s + \beta_1x + \beta_2x^2$$
Where $w$ is hourly wage, $w_0$ is the wage level of an individual with 0 years of schooling and work experience, $s$ is years of schooling, $x$ is work experience, and $\rho$, $\beta_1$, and $\beta_2$ are coefficients that can be viewed as the rates of return to schooling and work experience.

Numerous researchers have since tailored the Mincer equation by incorporating potential confounders, ensuring the mitigation of biased outcomes. A seminal study analyzing Chinese educational returns used the function:
$$ln(Y) = α_0 +α_1Edu +α_2Exper +α_3Exper^2+∑λ_jX+μ$$
Which adds $X$ that represents control variables such as age, gender, and geographical location [@Zhang2006]. Analogous modifications can be observed globally, although with variable inclusions that reflect regional societal and cultural differences. For example, an Indian study choose to include caste, relationship to the head of household, and the ability to speak and write English in the equation as control variables [@Bhandari2006]. This research also leaned on a modified version of the Mincer Earnings Function for its calculations on educational returns.

## Model set-up

### Multivariate regression models

Employing PSM-matched subsets, the educational returns for different family income quartile groups were computed using the modified Mincer Function:
$$ ln(w) = α_0 +α_1Edu +α_2Exp +α_3Exp^2 +α_4Gender \\ +α_5Hukou +α_6Marriage +α_7CCP +α_8East +μ$$
With $w$ representing hourly wage, $α_1$ representing returns to education and other coefficients representing the effects of previously identified covariates on the natural log of income.

Next, PSM-matched subset pairs were separated by family income quartiles and regressed independently to identify the differences between educational returns and the coefficients of covariates.

Finally, the Delta coefficient $\alpha_1$ from all paired comparisons across every CFPS iteration were compiled together. 5% t-tests were carried out on the data from each iteration to discern if the differences in educational returns between quartiles were statistically significant.

```{r}
#| echo: false
#| message: false
#| warning: false

# Regional polls analysis
trump_2024_regional <- trump_2024_lower %>% filter(state != "National")
harris_2024_regional <- harris_2024_lower %>% filter(state != "National")

trump_Pennsylvania <- trump_2024_regional %>% filter(state == "Pennsylvania")
trump_Georgia <- trump_2024_regional %>% filter(state == "Georgia")
trump_North_Carolina <- trump_2024_regional %>% filter(state == "North Carolina")
trump_Michigan <- trump_2024_regional %>% filter(state == "Michigan")
trump_Arizona <- trump_2024_regional %>% filter(state == "Arizona")
trump_Wisconsin <- trump_2024_regional %>% filter(state == "Wisconsin")
trump_Nevada <- trump_2024_regional %>% filter(state == "Nevada")

harris_Pennsylvania <- harris_2024_regional %>% filter(state == "Pennsylvania")
harris_Georgia <- harris_2024_regional %>% filter(state == "Georgia")
harris_North_Carolina <- harris_2024_regional %>% filter(state == "North Carolina")
harris_Michigan <- harris_2024_regional %>% filter(state == "Michigan")
harris_Arizona <- harris_2024_regional %>% filter(state == "Arizona")
harris_Wisconsin <- harris_2024_regional %>% filter(state == "Wisconsin")
harris_Nevada <- harris_2024_regional %>% filter(state == "Nevada")

combined_Pennsylvania <- full_join(
  trump_Pennsylvania,
  harris_Pennsylvania,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Georgia <- full_join(
  trump_Georgia,
  harris_Georgia,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_North_Carolina <- full_join(
  trump_North_Carolina,
  harris_North_Carolina,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Michigan <- full_join(
  trump_Michigan,
  harris_Michigan,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Arizona <- full_join(
  trump_Arizona,
  harris_Arizona,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Wisconsin <- full_join(
  trump_Wisconsin,
  harris_Wisconsin,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Nevada <- full_join(
  trump_Nevada,
  harris_Nevada,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

# mutate harris_win variable by comparing polled support rates
# preparing data for logistic regression
combined_Pennsylvania <- combined_Pennsylvania %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Georgia <- combined_Georgia %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_North_Carolina <- combined_North_Carolina %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Michigan <- combined_Michigan %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Arizona <- combined_Arizona %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Wisconsin <- combined_Wisconsin %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Nevada <- combined_Nevada %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

# Create 7 different logistic regressions to predict Harris's win rate
# in each state
logistic_Pennsylvania <- readRDS(here::here("models/Logistic_model_Pennsylvania.rds"))

logistic_Georgia <- readRDS(here::here("models/Logistic_model_Georgia.rds"))

logistic_North_Carolina <- readRDS(here::here("models/Logistic_model_North_Carolina.rds"))

logistic_Michigan <- readRDS(here::here("models/Logistic_model_Michigan.rds"))

logistic_Arizona <- readRDS(here::here("models/Logistic_model_Arizona.rds"))

logistic_Wisconsin <- readRDS(here::here("models/Logistic_model_Wisconsin.rds"))

logistic_Nevada <- readRDS(here::here("models/Logistic_model_Nevada.rds"))

# test logistic models for state specific analysis
# mutate fitted values

combined_Pennsylvania <- combined_Pennsylvania %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Pennsylvania, type = "response"), 2)
  )

roc_obj_Pennsylvania <- roc(
  combined_Pennsylvania$harris_win,
  combined_Pennsylvania$harris_win_prob
)

auc_value_Pennsylvania <- auc(roc_obj_Pennsylvania)

combined_Georgia <- combined_Georgia %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Georgia, type = "response"), 2)
  )

roc_obj_Georgia <- roc(
  combined_Georgia$harris_win,
  combined_Georgia$harris_win_prob
)

auc_value_Georgia <- auc(roc_obj_Georgia)

combined_North_Carolina <- combined_North_Carolina %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_North_Carolina, type = "response"), 2)
  )

roc_obj_North_Carolina <- roc(
  combined_North_Carolina$harris_win,
  combined_North_Carolina$harris_win_prob
)

auc_value_North_Carolina <- auc(roc_obj_North_Carolina)

combined_Michigan <- combined_Michigan %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Michigan, type = "response"), 2)
  )

roc_obj_Michigan <- roc(
  combined_Michigan$harris_win,
  combined_Michigan$harris_win_prob
)

auc_value_Michigan <- auc(roc_obj_Michigan)

combined_Arizona <- combined_Arizona %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Arizona, type = "response"), 2)
  )

roc_obj_Arizona <- roc(
  combined_Arizona$harris_win,
  combined_Arizona$harris_win_prob
)

auc_value_Arizona <- auc(roc_obj_Arizona)

combined_Wisconsin <- combined_Wisconsin %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Wisconsin, type = "response"), 2)
  )

roc_obj_Wisconsin <- roc(
  combined_Wisconsin$harris_win,
  combined_Wisconsin$harris_win_prob
)

auc_value_Wisconsin <- auc(roc_obj_Wisconsin)

combined_Nevada <- combined_Nevada %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Nevada, type = "response"), 2)
  )

roc_obj_Nevada <- roc(
  combined_Nevada$harris_win,
  combined_Nevada$harris_win_prob
)

auc_value_Nevada <- auc(roc_obj_Nevada)
```

We identified the AUC values and ROC plots for each of the seven swing-state logistic regressions. The ROC plots are presented in the Additional Data Details section (@sec-model-details), from @fig-pen to @fig-nev. The AUC values are: 0.90 for Pennsylvania, 0.88 for Georgia, 0.74 for North Carolina, 0.93 for Michigan, 1.00 for Arizona, 0.86 for Wisconsin, and 0.65 for Nevada. All states other than Nevada have acceptable or excellent AUC values, indicating that the respective logistic models are reliable when distinguishing cases where Harris wins. The low AUC value for Nevada is potentially caused by the limited sample size; there are only 9 polls conducted in Nevada.

Moreover, the sample size restrictions were the reason why we did not use training and testing data to validate our logistic models. The state with the greatest number of polls, Pennsylvania, only records 26 polls. Utilizing training and testing datasets implies that the datasets may include only a couple of data points in some states, which is insufficient to construct or test the model. On top of that, according to [@trainingerror], achieving low training error, such as by using a training-MSE-minimising regression model such as our own, is sufficient to ensure the predictive capabilities of a model bounded by some constant variance term.

```{r}
#| echo: false
#| message: false
#| warning: false

# find biases of pollsters from 2020 national polls data
# filter out national polls from the datasets
trump_2024_national <- trump_2024 %>% filter(state == "National")
harris_2024_national <- harris_2024 %>% filter(state == "National")
trump_2020_national <- trump_2020 %>% filter(state == "National")
biden_2020_national <- biden_2020 %>% filter(state == "National")

# Define actual results
actual_results_2020 <- list(Trump_pct = 47, Biden_pct = 51)

# Calculate bias per pollster
rep_pollster_bias <- trump_2020_national %>%
  group_by(pollster) %>%
  summarize(
    Trump_bias = mean(pct - actual_results_2020$Trump_pct, na.rm = TRUE)
  )

dem_pollster_bias <- biden_2020_national %>%
  group_by(pollster) %>%
  summarize(
    Biden_bias = mean(pct - actual_results_2020$Biden_pct, na.rm = TRUE)
  )

# adjust 2024 national polls percentages by the average pollster biases
# also mutate has_sponsor variable into dummy variable and remove NA's
trump_2024_national_adj <- trump_2024_national %>%
  left_join(rep_pollster_bias, by = "pollster") %>%
  mutate(
    pct_adj = pct - ifelse(!is.na(Trump_bias), Trump_bias, 0)
  ) %>%
  select(-Trump_bias)

# Adjust Harris 2024 National Polls
harris_2024_national_adj <- harris_2024_national %>%
  left_join(dem_pollster_bias, by = "pollster") %>%
  mutate(
    pct_adj = pct - ifelse(!is.na(Biden_bias), Biden_bias, 0)
  ) %>%
  select(-Biden_bias)

# model adjusted support percentage by time and control variables
# first split training and testing dataset using an 80/20 split
set.seed(11451)
trainIndex_trump_adj <- createDataPartition(trump_2024_national_adj$pct_adj,
  p = 0.8, list = FALSE
)
data_train_trump_adj <- trump_2024_national_adj[trainIndex_trump_adj, ]
data_test_trump_adj <- trump_2024_national_adj[-trainIndex_trump_adj, ]

trainIndex_harris_adj <- createDataPartition(harris_2024_national_adj$pct_adj,
  p = 0.8, list = FALSE
)
data_train_harris_adj <- harris_2024_national_adj[trainIndex_harris_adj, ]
data_test_harris_adj <- harris_2024_national_adj[-trainIndex_harris_adj, ]

model_date_harris_adj <- lm(pct_adj ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_harris_adj)
model_date_trump_adj <- lm(pct_adj ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_trump_adj)

# repeat for unadjusted support levels
trainIndex_trump <- createDataPartition(trump_2024_national$pct,
  p = 0.8, list = FALSE
)
data_train_trump <- trump_2024_national[trainIndex_trump, ]
data_test_trump <- trump_2024_national[-trainIndex_trump, ]

trainIndex_harris <- createDataPartition(harris_2024_national$pct,
  p = 0.8, list = FALSE
)
data_train_harris <- harris_2024_national[trainIndex_harris, ]
data_test_harris <- harris_2024_national[-trainIndex_harris, ]

model_date_harris <- lm(pct ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_harris)
model_date_trump <- lm(pct ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_trump)

# Functions to evaluate model using testing dataset
evaluate_model_adj <- function(model, test_data) {
  predictions <- predict(model, newdata = test_data)
  actuals <- test_data$pct_adj
  rmse <- sqrt(mean((predictions - actuals)^2, na.rm = TRUE))
  r_squared <- summary(model)$r.squared
  return(list(RMSE = rmse, R_squared = r_squared))
}

evaluate_model <- function(model, test_data) {
  predictions <- predict(model, newdata = test_data)
  actuals <- test_data$pct
  rmse <- sqrt(mean((predictions - actuals)^2, na.rm = TRUE))
  r_squared <- summary(model)$r.squared
  return(list(RMSE = rmse, R_squared = r_squared))
}

# Evaluate Harris model with adjusted support percentages
harris_evaluation_adj <- evaluate_model_adj(model_date_harris_adj, data_test_harris_adj)

# Evaluate Trump model with adjusted support percentages
trump_evaluation_adj <- evaluate_model_adj(model_date_trump_adj, data_test_trump_adj)

# Evaluate Harris model with unadjusted support percentages
harris_evaluation <- evaluate_model(model_date_harris, data_test_harris)

# Evaluate Trump model with unadjusted support percentages
trump_evaluation <- evaluate_model(model_date_trump, data_test_trump)

# Augment data with model predictions
model_harris_adj <- readRDS(here::here("models/MLR_adjusted_harris.rds"))
model_trump_adj <- readRDS(here::here("models/MLR_adjusted_trump.rds"))

harris_2024_national_adj <- harris_2024_national_adj %>% mutate(
  fitted_date = predict(model_harris_adj)
)

trump_2024_national_adj <- trump_2024_national_adj %>% mutate(
  fitted_date = predict(model_trump_adj)
)

# Unadjusted datasets
model_harris <- readRDS(here::here("models/MLR_harris.rds"))
model_trump <- readRDS(here::here("models/MLR_trump.rds"))

harris_2024_national <- harris_2024_national %>% mutate(
  fitted_date = predict(model_harris)
)

trump_2024_national <- trump_2024_national %>% mutate(
  fitted_date = predict(model_trump)
)

# evaluate models without using training and testing data
harris_eval_adj <- evaluate_model_adj(model_harris_adj, harris_2024_national_adj)

harris_eval <- evaluate_model(model_harris, harris_2024_national)

trump_eval_adj <- evaluate_model_adj(model_trump_adj, trump_2024_national_adj)

trump_eval <- evaluate_model(model_trump, trump_2024_national)

#### Logistic regression for national support ####

combined_national <- full_join(
  trump_2024_national,
  harris_2024_national,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_national <- combined_national %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

# repeat for adjusted support levels
combined_national_adj <- full_join(
  trump_2024_national_adj,
  harris_2024_national_adj,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_national_adj <- combined_national_adj %>% mutate(
  harris_win = case_when(
    pct_adj.y >= pct_adj.x ~ 1,
    pct_adj.y < pct_adj.x ~ 0
  )
)

# logistic model to predict the popular vote
model_logistic <- readRDS(here::here("models/Logistic_model.rds"))

model_logistic_adj <- readRDS(here::here("models/Logistic_model_adjusted.rds"))

# include predicted outcomes to the dataframes
combined_national <- combined_national %>%
  mutate(
    harris_win_prob = round(100 * predict(model_logistic, type = "response"), 2)
  )

combined_national_adj <- combined_national_adj %>%
  mutate(
    harris_win_prob = round(100 * predict(model_logistic_adj,
      type = "response"
    ), 2)
  )

#### Evaluate the Model ####

# ROC Curve and AUC
roc_obj <- roc(
  combined_national$harris_win,
  combined_national$harris_win_prob
)
auc_value <- auc(roc_obj)

roc_obj_adj <- roc(
  combined_national_adj$harris_win,
  combined_national_adj$harris_win_prob
)
auc_value_adj <- auc(roc_obj_adj)
```

For the analysis of the national popular vote, the logistic model of adjusted polls possess an AUC value of 0.98, and the logistic model of unadjusted polls have an AUC of 0.94, indicating high reliability of categorization. Their respective ROC plots are presented as @fig-log and @fig-logadj.

RMSE values and R squared values were calculated for the multiple linear regression models predicting popular vote. The RMSE values for all four models, adjusted and unadjusted support levels of the two candidates, are greater than 1, indicating that the fitted values and the support levels of actual polls are more than 1 percent apart on average, which is significant when analyzing popular vote. Moreover, all R squared values other than the model for the adjusted support rate of Trump, are smaller than 0.3, indicating the inability of the models to account for the variance of national support rates in the data. The model for the adjusted support rate of Trump has a R squared value of 0.69, indicating a relatively high proportion of variance in polled support rates being explained by the multiple linear regression model. This was expected, as the predictors such as a poll's end date and transparency does not necessarily correlate with the outcome of the polls. More predictive variables used by FiveThirtyEight [@5383] for their predictions are unavailable in the raw dataset, which is a key limitation of the data.

# Results {#sec-results}

## Swing states

@tbl-result shows the average likelihood percentage of Harris winning each swing state, calculated by using the logistic model on 1,000 hypothetical data points for each state, generated with the normal distribution of predictors and setting the end_date to election day. It indicates that Harris has a greater probability to win in Pennsylvania, Georgia, and Nevada, while Trump is more likely to win in North Carolina, Michigan, Arizona, and Wisconsin. The electoral votes in each state is also presented.

```{r}
#| echo: false
#| message: false
#| warning: false

#### Results ####
# To predict the outcome on election day, generate 1000 hypothetical data points
# using the election day as end_date and the normal curve of other input
# variables, mimicing all the situations that may happen on election day
hypothetical_data_Pennsylvania <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Pennsylvania$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Pennsylvania$transparency_score),
    sd = sd(combined_Pennsylvania$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Pennsylvania$sample_size),
    sd = sd(combined_Pennsylvania$sample_size)
  )
)

hypothetical_data_Georgia <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Georgia$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Georgia$transparency_score),
    sd = sd(combined_Georgia$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Georgia$sample_size),
    sd = sd(combined_Georgia$sample_size)
  )
)

hypothetical_data_North_Carolina <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_North_Carolina$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_North_Carolina$transparency_score),
    sd = sd(combined_North_Carolina$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_North_Carolina$sample_size),
    sd = sd(combined_North_Carolina$sample_size)
  )
)

hypothetical_data_Michigan <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Michigan$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Michigan$transparency_score),
    sd = sd(combined_Michigan$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Michigan$sample_size),
    sd = sd(combined_Michigan$sample_size)
  )
)

hypothetical_data_Arizona <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Arizona$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Arizona$transparency_score),
    sd = sd(combined_Arizona$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Arizona$sample_size),
    sd = sd(combined_Arizona$sample_size)
  )
)

hypothetical_data_Wisconsin <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Wisconsin$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Wisconsin$transparency_score),
    sd = sd(combined_Wisconsin$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Wisconsin$sample_size),
    sd = sd(combined_Wisconsin$sample_size)
  )
)

hypothetical_data_Nevada <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Nevada$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Nevada$transparency_score),
    sd = sd(combined_Nevada$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Nevada$sample_size),
    sd = sd(combined_Nevada$sample_size)
  )
)

# run the corresponding logistic regression on the hypothetical data of each
# state and average the outcomes to find the expected win rate of Harris
# on election day
hypothetical_data_Pennsylvania <- hypothetical_data_Pennsylvania %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Pennsylvania, newdata = ., type = "response"), 2)
  )

hypothetical_data_Georgia <- hypothetical_data_Georgia %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Georgia, newdata = ., type = "response"), 2)
  )

hypothetical_data_North_Carolina <- hypothetical_data_North_Carolina %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_North_Carolina, newdata = ., type = "response"), 2)
  )

hypothetical_data_Michigan <- hypothetical_data_Michigan %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Michigan, newdata = ., type = "response"), 2)
  )

hypothetical_data_Arizona <- hypothetical_data_Arizona %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Arizona, newdata = ., type = "response"), 2)
  )

hypothetical_data_Wisconsin <- hypothetical_data_Wisconsin %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Wisconsin, newdata = ., type = "response"), 2)
  )

hypothetical_data_Nevada <- hypothetical_data_Nevada %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Nevada, newdata = ., type = "response"), 2)
  )

# produce a table of mean winning probabilities
mean_probabilities <- tibble(
  State = c("Pennsylvania", "Georgia", "North Carolina", "Michigan", "Arizona", "Wisconsin", "Nevada"),
  Harris_Win_Probability = c(
    round(mean(hypothetical_data_Pennsylvania$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Georgia$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_North_Carolina$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Michigan$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Arizona$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Wisconsin$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Nevada$harris_win_prob, na.rm = TRUE), 2)
  )
)

# include electoral votes to determine who is in the lead
state_evs <- tibble::tribble(
  ~State, ~electoral_votes,
  "Alaska", 3,
  "Arizona", 11,
  "Arkansas", 6,
  "California", 55,
  "Colorado", 9,
  "Connecticut", 7,
  "Florida", 29,
  "Georgia", 16,
  "Illinois", 20,
  "Indiana", 11,
  "Iowa", 6,
  "Kansas", 6,
  "Louisiana", 8,
  "Maine", 2,
  "Maine CD-1", 1,
  "Maine CD-2", 1,
  "Maryland", 10,
  "Massachusetts", 11,
  "Michigan", 16,
  "Minnesota", 10,
  "Mississippi", 6,
  "Missouri", 10,
  "Montana", 3,
  "Nebraska", 2,
  "Nebraska CD-2", 1,
  "Nevada", 6,
  "New Hampshire", 4,
  "New Mexico", 5,
  "New York", 29,
  "North Carolina", 15,
  "Ohio", 18,
  "Oklahoma", 7,
  "Oregon", 7,
  "Pennsylvania", 20,
  "Rhode Island", 4,
  "South Carolina", 9,
  "South Dakota", 3,
  "Texas", 38,
  "Utah", 6,
  "Vermont", 3,
  "Virginia", 13,
  "Washington", 12,
  "Wisconsin", 10
)

mean_probabilities <- mean_probabilities %>%
  left_join(state_evs, by = "State")

#### Overall Probability of Winning ####
# Harris
# Create data frame of states
states <- data.frame(
  state = c("PA", "GA", "NC", "MI", "WI", "NV", "AZ"),
  prob = c(0.5551, 0.6801, 0.1033, 0.5151, 0.3792, 0.7679, 0),
  votes = c(20, 16, 15, 16, 10, 6, 11)
)

# Function to calculate probability of a specific combination
calc_combo_prob <- function(combo, states) {
  prob <- 1
  for (i in 1:nrow(states)) {
    if (i %in% combo) {
      prob <- prob * states$prob[i]
    } else {
      prob <- prob * (1 - states$prob[i])
    }
  }
  return(prob)
}

# Generate all possible combinations
n_states <- nrow(states)
total_prob <- 0

# Check each possible combination
for (i in 1:(2^n_states - 1)) {
  # Convert number to binary to get combination
  combo <- which(intToBits(i)[1:n_states] == 1)

  # Calculate total electoral votes for this combination
  votes <- sum(states$votes[combo])

  # If this combination has enough votes, add its probability
  if (votes >= 44) {
    prob <- calc_combo_prob(combo, states)
    total_prob <- total_prob + prob
  }
}

# Trump
# Create data frame of states
states <- data.frame(
  state = c("PA", "GA", "NC", "MI", "WI", "NV", "AZ"),
  prob = c(1 - 0.5551, 1 - 0.6801, 1 - 0.1033, 1 - 0.5151, 1 - 0.3792, 1 - 0.7679, 1 - 0), # Using 1-p for Trump's probabilities
  votes = c(20, 16, 15, 16, 10, 6, 11)
)

# Function to calculate probability of a specific combination
calc_combo_prob <- function(combo, states) {
  prob <- 1
  for (i in 1:nrow(states)) {
    if (i %in% combo) {
      prob <- prob * states$prob[i]
    } else {
      prob <- prob * (1 - states$prob[i])
    }
  }
  return(prob)
}

# Generate all possible combinations
n_states <- nrow(states)
total_prob <- 0

# Check each possible combination
for (i in 1:(2^n_states - 1)) {
  # Convert number to binary to get combination
  combo <- which(intToBits(i)[1:n_states] == 1)

  # Calculate total electoral votes for this combination
  votes <- sum(states$votes[combo])

  # If this combination has enough votes, add its probability
  if (votes >= 51) {
    prob <- calc_combo_prob(combo, states)
    total_prob <- total_prob + prob
  }
}
```

```{r}
#| label: tbl-result
#| tbl-cap: Predicted Harris Win Percentage in Swing States
#| echo: false
#| message: false
#| warning: false

mean_probabilities %>%
  gt() %>%
  fmt_number(
    columns = vars(Harris_Win_Probability),
    decimals = 2
  ) %>%
  cols_label(
    State = "State",
    Harris_Win_Probability = "Harris Win Probability (%)",
    electoral_votes = "Electoral Votes"
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 12,
    table.border.top.width = px(2),
    table.border.bottom.width = px(2)
  )
```

Combining with the fact that Harris requires 44 electoral votes to win and Trump requires 51, we have found using the probability function that Harris has a 38.3 chance to acquire more than 44 electoral votes, according to the probabilities and electoral votes in @tbl-result. On the other hand, Trump has a 61.7 probability to acquire more than 51 votes. Thus, our conclusion is that while both candidates possess a significant chance of winning, Trump has the greater probability of acquiring the necessary electoral votes from the seven swing states on election day.

## Popular vote

@fig-lradj presents the fitted values of the candidates' adjusted national support rates as well as the data points of polled support rates adjusted for historical pollster bias. It indicates that Trump is expected to have higher national support levels, but the two converges as election day approaches. This result is counter intuitive, as most national polls indicate that Harris will have the popular vote. However, the adjusted values for Trump's support level are significantly greater than the polled values because historically, the actual percentage of the popular vote for Trump in 2020 was significantly greater than his polled percentages of the popular vote. Thus, assuming that the 2024 national polls are equally biased against Trump compared to the 2020 polls, it is difficult to tell who will get the popular vote on election day. However, as previously stated, this assumption is likely false, as the candidates have changed.

```{r}
#| label: fig-lradj
#| fig-cap: Adjusted Popular Vote Prediction, 2024
#| echo: false
#| message: false
#| warning: false
# plot model predictions
# combine the datasets to create a single visual
combined_data_adj <- bind_rows(harris_2024_national_adj, trump_2024_national_adj)
combined_data <- bind_rows(harris_2024_national, trump_2024_national)

# create the plot that contains scattered points of adjusted percentages
# and lines for fitted values
ggplot(combined_data_adj, aes(x = end_date, y = pct_adj, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

\FloatBarrier

@fig-lr Presents the predicted unadjusted national support rates for Harris and Trump. The conclusions one can make from this plot is similar to those of most polls: Harris is expected to win the popular vote on any day, including the election day.

```{r}
#| label: fig-lr
#| fig-cap: Unadjusted Popular Vote Prediction, 2024
#| echo: false
#| message: false
#| warning: false
# Unadjusted support percentages
ggplot(combined_data, aes(x = end_date, y = pct, color = candidate_name)) +
  # Scatter points for actual adjusted percentages
  geom_point(alpha = 0.6) +
  # Regression lines
  geom_line(aes(y = fitted_date), size = 0.5) +
  xlim(as.Date("2024-07-19"), as.Date("2024-10-25")) +
  # Assign specific colors
  scale_color_manual(
    values =
      c("Kamala Harris" = "blue", "Donald Trump" = "red")
  ) +
  # Customize the plot appearance
  theme_classic() +
  labs(
    y = "Adjusted Percent Support",
    x = "End Date",
    color = "Candidate"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

\FloatBarrier

The regression coefficients for the logistic and linear models are presented in the Additional Data Details section (@sec-model-details) because they are not the outcome of interest.

# Discussion {#sec-discussion}

## The illusion of polarization

As mentioned in the introduction, one thing that stands out about the 2024 election in particular is just how polarized the two factions are. Instead of merely being a choice over which individual to helm the country over the next four years, this particular election is being framed more and more as a “fight to protect our democracy” from both sides of the debate stand. This, while evidently a new level of polarization, merely mirrors the trend in recent years more than anything else: party loyalty has been on a steady uprise since 1972, with loyalty in weak partisan and leaning independent voters increasing by approximately 50% in the 40 years that followed [@partisanship]. New records were further set in 2016, with Republican validated voters voting Trump a staggering 92% to 4% for Clinton, and Democratic validated voters voting Clinton an equally extreme 94% to 5% Trump support [@pew2018voters].

New voting statistics, however, suggest that this polarization is likely only more prominent in the upper echelons of political representation, however: a Washington Post article found that approximately 1 in 8 women and 1 in 10 men are now voting differently from their spouses compared to before [@wapo2024women], a development that suggests a decrease in long-increasing partisanship trends. This is especially pronounced since households in the States overwhelmingly vote together: at least 70% of them do, at least, which means that this value is projected to decrease from 70% to approximately 62% [@households]. In short: while it seems that ideologies are shifting more and more away from each other, the stats say otherwise. So does the Overton window [@overton].

## Win probabilities versus support ratings: why our predictions are as extreme as they are

Our model led to a win probability prediction with numbers much more skewed towards one side than most pollsters currently predict: under this model, we find that Harris has a 40-60 chance against Trump, whereas most pollsters nowadays hover somewhere between 50-50 to 45-55 (towards either side). This is mainly a consequence of using a multinomial logistic regression model as opposed than other models – instead of predicting the support rates of the two candidates, such a model outputs the probability that a random support rate value from the predicted distribution of Harris support rates is larger than that of the predicted distribution of Trump support rates. This means that in states such as Nevada, where 538 gives Trump a +3 advantage, factoring in for margin of error and standard error means that the probability of Harris winning the state is not in fact 47% but rather much closer to zero, which we decided is a more intuitive way of understanding election odds.

A final reason for this is how swing states are polled much more sample-size-wise than other states, leading to a much higher power value than polls from said other states – a +3 lead in a sample of 1000 people from California’s 39 million might not mean too much, but a +3 lead in a sample of 3000 people from Nevada’s 3 million represents the actual results for 0.3% of the state’s population, holding much more relative weight when it comes to election predictions.

## Weaknesses and next steps

The dataset used for this analysis has a number of clear limitations. First of all, there is a lack of high-quality polls. Though FiveThirtyEight distributes each pollster with a rating out of 3.0, almost all pollsters under the rating of 2.7 produces outcomes that are unreliable and highly volatile. For this research, it was only possible for us to use pollsters with a 3.0 rating when analyzing national polls; we had to lower the standards to 2.5 for the analysis of swing states, as the number of high-quality state-specific polls are very limited. The inclusion of unreliable data reduces the predictability of our logistic models and thus our final outcome. The lack of predictability of our Nevada model is a perfect example of how low-quality data and the lack of data points impacts our ability to create effective models. However, we acknowledge that this is beyond the control of FiveThirtyEight, as they do not conduct the polls.

The second major challenge presented by the dataset is the lack of predictors. FiveThirtyEight utilizes predictive variables that are unavailable to us, such as weighted values that represent the extent to which a pollster is partisan [@5383]. The predictors used in this research, such as the transparency score of the pollster, are not related to their bias towards any candidate from a partisanship point of view, limiting the predictive power of our models.

Finally, the lack of data regarding the Harris-Trump race, due to Harris starting her presidential race uncharacteristically late, is also a main reason for the limitations of our model.

Potential next steps could be taken to improve the model in the future: more data could be aggregated manually (from pollsters not recorded by FiveThirtyEight), stronger predictor variables that are specifically more correlated with partisanship could be marked and utilised in our model, and individual weightings could be utilised and applied to the data based on each pollster's "track record" with regards to the error, bias, and variance in their previous polls and estimates.

\newpage

\appendix

# Appendix {.unnumbered}

## Pollster methodology overview and evaluation

The Quinnipiac University Poll conducts independent polling in swing states. To analyze their methodology, we look specifically into their October 2024 Pennsylvania polls [@quinnipiac].

Their target population is likely voters aged 18 and older in Pennsylvania [@quinnipiacmethodology]. To reach this population, they use likely voters aged 18 and older with phone numbers (both landline and cell) as their sampling frame, i.e. the frame of possible subjects that they sample observations from. Quinnipiac University employed Random Digit Dialing (RDD) to generate their sample of 2,186 respondents. This dual-frame approach reflects modern communication patterns, with 1,644 cell phone and 542 landline completions. However, using phones as a sampling frame means they cannot reach voters without phone access, introducing potential coverage bias.

Their sampling approach uses stratification by Census division according to area code, meaning that they divide Pennsylvania into geographic regions before using RDD to sample within each region. This strategy ensures even geographic representation but adds complexity to the sampling process [@pollofpolls]. For each selected number, they attempt contact at least three times before marking it as non-responsive. For landline calls, they ask to speak with the household member who has the next birthday, a simple but effective randomization technique. Afterwards, a series of screening questions confirm that the subject is indeed a likely voter, after which the subject's responses are then formally taken as part of the sample.

After collecting responses, Quinnipiac adjusts their data through post-stratification weighting. In this weighting, they compare their sample's demographic composition to known population benchmarks from the Census (like age, gender, education, and race distributions in Pennsylvania) and adjust the weight given to each response to match these benchmarks. For example, if their sample has too few young voters compared to Census data, responses from young voters would be weighted more heavily. While this helps correct for sampling imbalances, it can increase the variance in their estimates if the weights vary substantially [@poststratification].

The survey administration addresses measurement issues through its design. Live interviewers conduct all interviews, enabling question clarification and generating higher response rates compared to automated systems. However, live interviewers may introduce social desirability bias, where respondents might modify their answers to appear more socially acceptable. This becomes particularly relevant in political polling, where respondents might hesitate to express unpopular political views [@politicalpolling].

Several types of bias affect the poll's results. Self-selection bias occurs because certain types of people (typically those more politically engaged or with stronger views) are more likely to agree to participate in the survey. Non-response bias arises when people who respond differ systematically from those who don't - for instance, busier people might be less likely to answer calls, potentially underrepresenting certain occupational groups [@nonresponse]. Coverage bias means some groups (like those without phones) have no chance of being included in the sample.

Ethics-wise, the survey does a good job of informing individuals of the details of the survey, such as its purpose and how participants' data is to be used. Personally-identifiable data (i.e. name, phone number, etc.) is not collected during the survey other than to ensure that repeat numbers are not drawn, and the usage of live interviewers to conduct the survey ensures that concerns about the ethics of the survey can be voiced and answered on-the-spot. Combined with the lack of financial incentive for participants, which suggests that all participants participate of their complete free will and intention, this is an ethically sound methodology for conducting a political survey.

Finally, the poll also faces common challenges in political polls such as this one - the five-day field period (October 24-28) may miss opinion changes close to election day, and while weighting adjustments help correct for demographic imbalances, they may increase variance in the estimates if some groups need to be weighted heavily to match population benchmarks, for example.

Overall, Quinnipiac's methodology represents a balanced approach to managing practical constraints and statistical rigor in modern political polling - while some common biases are still likely to skew the poll results off the true support levels for Harris, for example, the poll uses methods such as post-stratification weighting to tradeoff biases at the cost of model variance [@weighting]. Using live interviewers unavoidably introduces social desirability bias, however, and significantly increases poll costs per quota. Modifying the methodology to remove this aspect of the survey would potentially reduce bias and allow for larger samples to be taken, in turn opening up possibilities for cross-validation and the such, which then reduces the effect of increased model variance on the final results.

## Idealised methodology

With a \$100,000 budget, our approach focuses on producing accurate state-level estimates in key battleground states, which would then inform our national forecast. We prioritize Pennsylvania, Michigan, Wisconsin, Georgia, Arizona, and Nevada, allocating resources proportionally based on each state's electoral importance and expected margin of victory.

Our sampling strategy employs both probability and non-probability methods. Probability sampling (60% of budget, i.e. \$60k) means every member of our target population has a known, non-zero chance of being selected - this allows us to calculate proper margins of error and make statistical inferences about the population. For this, we use dual-frame random digit dialing (RDD) for phone surveys and address-based sampling (ABS) for mail-to-web recruitment. RDD involves generating random phone numbers within active area codes, while ABS uses the U.S. Postal Service's delivery database as a sampling frame, both proven methods of sampling [@cummings]. The ABS approach helps reach households without reliable phone access. We stratify our sample by geography, demographics, and previous voting patterns to ensure representation across key subgroups - meaning we divide the population into these subgroups and sample from each independently.

For non-probability sampling (40% of budget, i.e. \$40k), where respondents' selection probabilities are unknown and not everyone has a chance of being selected, we recruit through multiple online panel vendors and use targeted social media advertising to reach traditionally underrepresented groups. While this approach introduces potential selection bias because participants self-select into the sample, it helps reach younger voters who are less responsive to traditional survey methods [@nonprobabilitysampling]. We implement quota sampling within these non-probability samples to match key demographic targets - for example, stopping collection from certain demographic groups once their quota is filled.

Respondent recruitment uses multiple contact methods - mail, email, text, and phone - with attempts made at different times and days to maximize response rates. We offer a \$10 gift card incentive for completed surveys and provide both English and Spanish language options. This mixed-mode contact strategy helps reduce non-response bias by providing multiple ways to participate.

Data validation is crucial for maintaining quality. We cross-reference responses with voter files where available - meaning we check if respondents' self-reported registration status matches official records. We screen for duplicate responses using IP addresses and phone numbers, and implement attention checks within the survey (questions with known correct answers to ensure respondents are reading carefully). Speed checks identify rushed responses that might indicate low-quality data by flagging completions that fall below a minimum reasonable completion time, while consistency checks across related questions help identify potentially fraudulent responses by looking for logical contradictions in answers.

Our weighting approach uses post-stratification to known population benchmarks - this means we adjust the weight given to each response so that our sample matches known population characteristics. For example, if our sample has 30% college graduates but the population has 40%, we would give more weight to responses from college graduates. We include demographics (age, race, education, gender), geographic location, past voting behavior, and party registration in our weighting scheme. We produce daily estimates using a 7-day rolling average, which helps smooth out daily fluctuations while remaining responsive to real changes in voter preferences [@weighting2].

Finally, how ethics are handled is a crucial part of any survey methodology. In this idealised methodology, consent will be asked for at the beginning of the survey, full disclosure of how information is used will be given beforehand and no self-identifiable information will be recorded (so no names, phone numbers, etc.). The \$10 incentive is enough to hopefully make it wortile for participants' time, but also not ideally not significant enough of an incentive to make individuals suppress otherwise deal-breaking concerns with the survey purely for the sake of the incentive.

The survey instrument itself focuses on six key areas: screening questions to identify likely voters, voting intentions (including direct questions about Trump vs. Harris preferences), political preferences, demographics, issue priorities, and media consumption patterns. We've implemented this survey design in Google Forms, which can be found here: https://forms.gle/pk7vDiMHwEGLMK849

This methodology balances statistical rigor with practical constraints, while acknowledging and attempting to address the key challenges in modern political polling: declining response rates, coverage bias, and the increasing difficulty of reaching a representative sample of likely voters.

## Idealised survey

The survey, made using Google Forms, is linked here: https://forms.gle/pk7vDiMHwEGLMK849 Note that the questions are identical for both the phone and online surveys bar q6. A copy of the survey that is identical to the one implemented in the Google Forms above is presented below: Thank you for participating in this survey about the 2024 U.S. Presidential Election. This survey is part of a research project at the University of Toronto studying voting intentions and political attitudes.

Estimated completion time: 8-10 minutes

Your responses will be kept confidential and used only for research purposes. Email information, and any other information that may personally identify you, is not gathered. You may skip any questions you prefer not to answer, though complete responses are most helpful for our research.

For questions or concerns about this survey, please contact: andrew.goh\@mail.utoronto.ca

SCREENING SECTION: Q1. Are you 18 years of age or older?

Yes No \[END SURVEY\]

Q2. Are you a U.S. citizen?

Yes No \[END SURVEY\]

Q3. Are you registered to vote at your current address?

Yes No Not sure \[If No or Not sure: Do you plan to register before the November 2024 election?\]

VOTING INTENTION: Q4. How likely are you to vote in the 2024 presidential election?

Definitely will vote Probably will vote Might or might not vote Probably will not vote Definitely will not vote

Q5. If the 2024 presidential election were held today, and the candidates were Kamala Harris (Democrat) and Donald Trump (Republican), who would you vote for?

Kamala Harris Donald Trump Another candidate (please specify) Would not vote Not sure

ATTENTION CHECK: Q6. To ensure you're reading carefully, please select "Somewhat disagree" for this question: "I enjoy following political news."

Strongly agree Somewhat agree Somewhat disagree Strongly disagree No opinion

POLITICAL PREFERENCES: Q7. Generally speaking, you consider yourself a:

Democrat Republican Independent Something else (please specify)

Q8. How would you rate the current state of the U.S. economy?

Very poor Poor Fair Good Excellent

ISSUE PRIORITIES: Q9. Which ONE of the following issues is most important to you when deciding how to vote?

Economy and jobs Immigration Healthcare Climate change Crime and public safety Education National security Abortion rights Gun policy Something else (please specify)

For each of the following issues, please indicate whether you think Kamala Harris or Donald Trump would do a better job handling it:

Q10. Economy and jobs:

Kamala Harris would do better Donald Trump would do better No difference Not sure

Q11. Human rights and freedom of speech: \[Same options\] Q12. Abortion: \[Same options\] Q13. Healthcare: \[Same options\] Q14. Immigration: \[Same options\] Q15. National security: \[Same options\]

MEDIA CONSUMPTION: Q16. Where do you most often get your news about politics? (Select all that apply)

Network TV news (ABC, CBS, NBC) Cable TV news (CNN, Fox News, MSNBC) Local TV news Radio Print newspapers News websites Social media Friends and family Other (please specify)

Q17. How many hours per day do you typically spend following news about politics?

Less than 1 hour 1-2 hours 2-4 hours More than 4 hours

DEMOGRAPHICS: Q18. What is your age?

18-24 25-34 35-44 45-54 55-64 65 or older

Q19. What is your gender?

Male Female Non-binary/Other Prefer not to say

Q20. What is your race/ethnicity? (Select all that apply)

White Black or African American Hispanic or Latino Asian Native American Other (please specify) Prefer not to say

Q21. What is the highest level of education you have completed?

Less than high school High school graduate Some college Associate's degree Bachelor's degree Graduate degree Prefer not to say

Q22. What was your total household income before taxes in 2023?

Under \$25,000 \$25,000-\$49,999 \$50,000-\$74,999 \$75,000-\$99,999 \$100,000-\$149,999 \$150,000 or more Prefer not to say

CONSISTENCY CHECK: Q23. Looking ahead to November 2024, if Kamala Harris is the Democratic nominee and Donald Trump is the Republican nominee, how do you think you will vote?

Kamala Harris Donald Trump Another candidate (please specify) Would not vote Not sure

\[END OF SURVEY\] Thank you for completing this survey about the 2024 U.S. Presidential Election. Your responses will help us better understand voter preferences and political attitudes across the country. If you have any questions about this research or would like to be informed about the results, please contact andrew.goh\@mail.utoronto.ca. Your time and participation is greatly appreciated.

# Additional data details

# Model details {#sec-model-details}

## Diagnostics

```{r}
#| label: fig-pen
#| fig-cap: ROC Curve for Harris Win Prediction, Pennsylvania
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Pennsylvania)
```

```{r}
#| label: fig-geo
#| fig-cap: ROC Curve for Harris Win Prediction, Georgia
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Georgia)
```

```{r}
#| label: fig-nor
#| fig-cap: ROC Curve for Harris Win Prediction, North Carolina
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_North_Carolina)
```

```{r}
#| label: fig-mic
#| fig-cap: ROC Curve for Harris Win Prediction, Michigan
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Michigan)
```

```{r}
#| label: fig-ari
#| fig-cap: ROC Curve for Harris Win Prediction, Arizona
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Arizona)
```

```{r}
#| label: fig-wis
#| fig-cap: ROC Curve for Harris Win Prediction, Wisconsin
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Wisconsin)
```

```{r}
#| label: fig-nev
#| fig-cap: ROC Curve for Harris Win Prediction, Nevada
#| echo: false
#| message: false
#| warning: false

plot(roc_obj_Nevada)
```

```{r}
#| label: fig-log
#| fig-cap: ROC Curve for Popular Vote Prediction
#| echo: false
#| message: false
#| warning: false

plot(roc_obj)
```

```{r}
#| label: fig-logadj
#| fig-cap: ROC Curve for Popular Vote Prediction, Adjusted
#| echo: false
#| message: false
#| warning: false

plot(roc_obj)
```

\FloatBarrier

## Coefficients

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Pennsylvania <- tidy(logistic_Pennsylvania)

# Display the table using knitr::kable
kable(tidy_Pennsylvania,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Pennsylvania"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Georgia <- tidy(logistic_Georgia)

# Display the table using knitr::kable
kable(tidy_Georgia,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Georgia"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Michigan <- tidy(logistic_Michigan)

# Display the table using knitr::kable
kable(tidy_Michigan,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Michigan"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_North_Carolina <- tidy(logistic_North_Carolina)

# Display the table using knitr::kable
kable(tidy_North_Carolina,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for North_Carolina"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Arizona <- tidy(logistic_Arizona)

# Display the table using knitr::kable
kable(tidy_Arizona,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Arizona"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Wisconsin <- tidy(logistic_Wisconsin)

# Display the table using knitr::kable
kable(tidy_Wisconsin,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Wisconsin"
)
```

```{r}
#| echo: false
#| message: false
#| warning: false
# Tidy the logistic regression model
tidy_Nevada <- tidy(logistic_Nevada)

# Display the table using knitr::kable
kable(tidy_Nevada,
  digits = 2,
  col.names = c("Term", "Estimate", "Std. Error", "Statistic", "p-value"),
  caption = "Logistic Regression Coefficients for Nevada"
)
```

\newpage

# References